### 2025-05-18 mss.tex
Post

\documentclass[twoside]{article}

\usepackage{ustj}

\addbibresource{mss.bib}

\newcommand{\authorname}{N. E. Davis}
\newcommand{\authorpatp}{\patp{lagrev-nocfep}}
\newcommand{\affiliation}{Zorp Corp}

%  Make first page footer:
\fancypagestyle{firststyle}{%
\fancyhf{}% Clear header/footer
\fancyhead{}
\fancyfoot[L]{{\footnotesize
              %% We toggle between these:
              Manuscript submitted for review.\\
              % {\it Urbit Systems Technical Journal} II:1 (2025):  1–3. \\
              ~ \\
              Address author correspondence to \authorpatp.
              }}
}
%  Arrange subsequent pages:
\fancyhf{}
\fancyhead[LE]{{\urbitfont Urbit Systems Technical Journal}}
\fancyhead[RO]{A History of Nock}
\fancyfoot[LE,RO]{\thepage}

%%MANUSCRIPT
\title{Hot Nock:  A History of the Nock Combinator Calculus}
\author{\authorname~\authorpatp \\ \affiliation}
\date{}

\begin{document}

\maketitle
\thispagestyle{firststyle}

\begin{abstract}
\end{abstract}

% We will adjust page numbering in final editing.
\pagenumbering{arabic}
\setcounter{page}{1}

\tableofcontents

\section{Introduction}

Nock is a combinator calculus which serves as the computational specification layer for the Urbit and Nockchain/NockApp systems.  It is a hyper-\textsc{risc} instruction set architecture (ISA) intended for execution by a virtual machine (but see \citet{Mopfel2025}, pp.~XX–XX herein).  Nock's simplicity and unity of expression make it amenable to proof-based reasoning and guarantees of correctness.  Its Lisp-like nature surfaces the ability to introspect on the code itself, a property which higher-level languages compiling to it can exploit.  Nock permits itself a finite number of specification changes, called ``decrements'' or ``kelvins'', which allow it to converge on a balance of expressiveness and efficacy.  This article outlines the extant historical versions of the Nock combinator calculus and reconstructs the motivation for the changes made at each kelvin decrement.  It begins with an exposition of Nock as a tool of computation, outlines the history of Nock's decrements, and speculates on motivations for possible future developments.

\section{Nock as a Combinator Calculus}

Fundamental computer science research has identified a family of universal computers which may be instantiated in a variety of ways, such as the Turing machine, the lambda calculus, and the combinator calculus.  Equivalence theorems such as the Church–Turing thesis show that these systems are equivalent in their computational power, and that they can be used to compute any computable function.  The combinator calculus is a family of systems which use a small set of combinators to express computation.  The most well-known member of this family is the \textsc{ski} combinator calculus, which uses only three combinators: \textsc{s}, \textsc{k}, and \textsc{i}.  Other members of this family include the \textsc{bckw} combinator calculus and the \textsc{h} combinator calculus.  These systems are all equivalent in their computational power, but they differ in their syntax and semantics.  The Nock combinator calculus is an extension of the \textsc{ski} combinator calculus which adds a few axiomatic rules to navigate and manipulate binary trees, carry out a very primitive arithmetic, and provide for side effects.

Perhaps better put, Nock is a family of combinator calculi that sequentially converge on an ``optimal'' expressiveness for certain design desiderata.  This includes an economy of expression (thus several macro opcodes) and consideration of how a higher-level language would invoke stored procedural expressions.  Furthermore an opcode exists which produces and then ignores a computation, intended to signal to a runtime layer that a side effect may be desired by the caller.

Nock bears the following characteristics:

\begin{itemize}
  \item  Turing-complete.  Put formally, Turing completeness (and thus the ability to evaluate anything we would call a computation) is exemplified by the $μ$-recursive functions \citep{WikipediaMuRecursive}.  In practice, these amount to operations for constant, increment, variable access, program concatenation, and looping \citep{Raphael2012}.  Nock supports these directly through its primitive opcodes.
  % WikipediaMuRecursive https://en.wikipedia.org/wiki/General_recursive_function
  % Raphael2012 https://cs.stackexchange.com/questions/991/are-there-minimum-criteria-for-a-programming-language-being-turing-complete

  \item  Functional (as in language).  Nock is a pure function of its arguments.  In practice, the Urbit operating system provides a simulated global scope for userspace applications, but this virtualized environment reduces to garden-variety Nock.  (See \citet{Davis2025b}, pp.~XX–XX in this volume, for details of a Nock virtualized interpreter.)
  \item  Subject-oriented.  Nock evaluation consists of a formula as a noun to be evaluated against a subject as a noun.  Taken together, these constitute the entire set of inputs to a pure function.

    \begin{quote}
    Everything about a scope, including name bindings, aliases, and docstrings, is stored in the subject's type. This allows Hoon's compilation discipline to be similarly minimal: the compiler is a function from subject type and Hoon source to product type and compiled Nock. Running this Nock against a value of the subject type produces a vase of the result. It's hard to imagine a more streamlined formalism for compilation.  \citep{Blackman2020}
    %(Ted Blackman, [“Why Hoon”](https://urbit.org/blog/why-hoon))
    \end{quote}

    \noindent
    Some Nock opcodes alter the subject, such as a variable declaration, by producing a new subject which is utilized for subsequent axis lookups.

  \item  Homoiconic.  Nock unifies code and data under a single representation.  A Nock atom is a natural number, and a Nock cell is a pair of nouns.  Every Nock noun is acyclic, and every Nock expression is a binary tree.  For example, Nock expressions intended to be evaluated as code are often pinned as data by the constant opcode until they are retrieved by evaluating the constant opcode at that axis.

  \item  Untyped.  Nock is untyped, meaning that it does not impose any type system on the expressions it evaluates.  Nock ``knows'' about the natural numbers in two senses:  such are used for addressing axes in the binary tree of a noun, and such are manipulated and compared using the increment and equality opcodes.

  \item  Solid-state.  A Nock interpreter is a solid-state machine, meaning that it operates from a state to a new state strictly according to inputs as a pure lifecycle function.  The Nock interpreter must commit the results of a successful computation as the new state before subsequent computations, or events, can be evaluated.  Transient evaluations (uncompleted events) and crashes (invalid evaluations) may be lost without consequence, and the Nock interpreter layer persists the underlying state of the machine.
\end{itemize}

We have asserted without demonstration that Nock is a combinator calculus.  We now show that this is the case, with reference to Nock 4K, the latest specification.  The simplest combinator calculus consists of only three combinators \citep{SomeoneSmart}, \textsc{s}, \textsc{k}, and \textsc{i}.  These combinators are defined as follows:

\begin{enumerate}
  \item  \textsc{s} substitution.  This corresponds to Nock 4K's opcode 2, which substitutes the second argument into the first argument at the third argument's axis.  (There are some subtle differences to Nock's expression of \textsc{s} as opcode 2 that we will elide as being fundamentally similar, but perhaps worthy of its own monograph.)
  \item  \textsc{k} constant.  This corresponds to Nock 4K's opcode 1, which yields its argument as a constant noun.
  \item  \textsc{i} identity.  This corresponds to Nock 4K's opcode 0, a generalized axis lookup operator, which can trivially retrieve the current subject or expression as well as any children.
\end{enumerate}

\noindent
While Nock introduces a few more primitive operations as a practicality, the above identities establish its bona fides as a combinator calculus capable of general computation.  Similar to Haskell Curry's \textsc{bckw} system, which can be written in forms reducible to \textsc{ski}, Nock provides a set of primitive rules and a set of economic extended rules for convenience in writing a compiler.  (See \citet{Galebach2025}, pp.~XX–XX in this volume, for exposition on how to evaluate a Nock expression by hand or by interpreter.)

### Event Loop

The beating heart of an operational system using Nock is the event loop formula.  For each event that is injected, the system runs this formula in the kernel's `+aeon` arm:

```
.*([arvo -.epic] [9 2 10 [6 0 3] 0 2])
```

Let's talk through what is going on here.

- The subject consists of a pair of `arvo`, the evolution from underlying operating system and standard library until the current epitome state, and the head of `epic`, where `epic` is the remaining event log to be evaluated.
- The formula is a 9 which runs the head of the event log, `[0 3]`, against the whole subject, then replaces that into the head of the subject.
- The next event is then evaluated, until the list is empty.
- When it reaches the end of the event log, the runtime holds it in stasis until the next event is prepared to be injected.

### Jets

Jet dispatch does incur some overhead, which is generally worthwhile

## Tree-Walking Interpreter

Nock is interpreted in the sense that it is produced to be executed by a virtual machine layer.  A human end-user language will be parsed and interpreted into an abstract syntax tree, then converted into raw Nock for evaluation.  This Nock is itself interpreted at runtime by a tree-walking interpreter.

This interpreter takes a Nock formula as input, recursively traverses the formula tree, evaluates each operation according to its subject, and returns the resulting noun.

The upside of Nock is that this process is fully general and portable.  And since Nock is a both an ISA and a specification, when jetted code is available it can be evaluated in a faster way as much as possible.

So, some of you are going to ask me, what's the alternative to a tree-walking interpreter?  To answer that, let's think about the Nock execution pattern.  The major downside of Nock as an ISA is that it is a hyper-RISC that does not closely map to von Neumann architecture CPU operations.  Modern computer architectures prefer to grab arrays in a hierarchy of caches, but a binary tree implementation does not lend itself well to this arrangement.  Furthermore, because Nock is homoiconic, it is not straightforward to tell from a tree-walking interpreter if a given noun is code or data.  A newer paradigm, called subject knowledge analysis, proposes to fix this by a system of JIT-like affordances produced by a static analysis at compile time.  SKA enables efficient analysis and execution of Nock code by unearthing the computational structure of the code even in an untyped procedure-free environment.

### Nock Bytecode

There are currently two practical Nock interpreters in production, Urbit's Vere and Zorp's Sword.  They both currently use a tree-walking interpreter, but a more sophisticated protocol is in active development.  We will take a brief look at Vere's bytecode implementation.

Vere walks the Nock tree to produce a linear bytecode array for rapid evaluation.  The bytecode interpreter is a stack machine built on top of an allocator.  It uses a threaded code pattern or a computed `GOTO`.  While it is highly optimized on its own terms, the gains are often clawed back by other parts of the system.

Vere can reveal the bytecode to us if we query it with an `%xray` Nock 11 hint.

For instance

We can see these by passing a Nock 11 `%xray` hint at the command line.

```
{[lils 65535] halt}
> ~>  %xray  65.535
65.535
{[libl i:0] halt}
> ~>  %xray  65.536
65.536

```

As with other bytecodes, these can be rather opaque.  (In this case, we are sometimes left with labels and pointers to data in other locations rather than the complete data set.)  Vere's bytecode interpreter is, I have been told, rather simple and primitive.

```hoon
> !=  +(41)
[4 1 41]

> ~>  %xray  +(41)
{[lilb 41] bump halt}

> !=  (dec 43)
[8 [9 2.398 0 1.023] 9 2 10 [6 7 [0 3] 1 43] 0 2]

> ~>  %xray  (dec 43)
{[fask 1023] [kicb 1] snol head swap tail [lilb 43] musm [ticb 0] halt}
```

\section{Nock's Decrements}

The earliest extant Nock is \texttt{U}, a proto-Nock posted to the \textit{Lambda the Ultimate} blog in 2006 \citep{Yarvin2006}.\footnote{Curtis Yarvin was consulted for elements of this history.  Unfortunately many elements of the original prehistory of Nock appear to be lost to the sands of time on unrecoverable hard drives.}  The draft is versioned 0.15, but it is unclear whether this relates to kelvin versioning or not.  The document is produced in full in Listing~\ref{lst:u}.
% Yarvin2006, http://lambda-the-ultimate.org/node/1269
% Yarvin2006a, https://web.archive.org/web/20060701070718/http://urbit.sourceforge.net/u.txt

Yarvin had likely been influenced by X.
(He later cited Y.)

\begin{lstlisting}[label=lst:u,caption={U: Definition.  The earliest extant Nock.  Only layout has been changed for print.},style=listingcode]
U: Definition

1 Purpose

    This document defines the U function and its data
    model.

2 License

    U is in the public domain.

3 Status

    This text is a DRAFT (version 0.15).

4 Data

    A value in U is called a "term."  There are three
    kinds of term: "number," "pair," and "foo."

    A number is any natural number (ie, nonnegative
    integer).

    A pair is an ordered pair of any two terms.

    There is only one foo.

5 Syntax

    U is a computational model, not a programming
    language.

    But a trivial ASCII syntax for terms is useful.

5.1 Trivial syntax: briefly

    Numbers are in decimal.  Pairs are in parentheses
    that nest to the right.  Foo is "~".

    Whitespace is space or newline.  Line comments
    use "#".

5.2 Trivial syntax: exactly

    term    : number
            | 40 ?white pair ?white 41
            | foo

    number  : 48
            | [49-57] *[48-57]

    pair    : term white term
            | term white pair

    foo     : 126

    white   : *(32 | 10 | (35 *[32-126] 10))

6 Semantics

    U is a pure function from term to term.

    This document completely defines U.  There is no
    compatible way to extend or revise U.

6.1 Rules

    [name]   [pattern]                [definition]

    (a)      ($a 0 $b)                $b
    (b)      ($a 1 $b $c)             1
    (c)      ($a 1 $b)                0
    (d)      ($a 2 0 $b $c)           $b
    (e)      ($a 2 %n $b $c)          $c
    (f)      ($a 3 $b $c)             =($b $c)
    (g)      ($a 4 %n)                +%n

    (h)      ($a 5 (~ ~ $b) $c)       $b
    (i)      ($a 5 (~ $b $c) $d)      *($a $b $c $d)
    (j)      ($a 5 (~ ~) $b)          ~
    (k)      ($a 5 (~ $b) $c)         *($a $b $c)
    (l)      ($a 5 ($b $c) $d)
                            (*($a $b $d) *($a $c $d))
    (m)      ($a 5 $b $c)             $b

    (n)      ($a 6 $b $c)   *($a *($a 5 $b $c))
    (o)      ($a 7 $b)                *($a 5 $a $a $b)
    (p)      ($a 8 $b $c $d)          >($b $c $d)

    (q)      ($a $b $c)     *($a 5 *($a 7 $b) $c)
    (r)      ($a $b)                  *($a $b)
    (s)      $a                       *$a

    The rule notation is a pseudocode, only used in
    this file. Its definition follows.

6.2 Rule pseudocode: briefly

    Each line is a pattern match.  "%" means
    "number."  Match in order.  See operators below.

6.3 Rule pseudocode: exactly

    Both pattern and definition use the same
    evaluation language, an extension of the trivial
    syntax.

    An evaluation is a tree in which each node is a
    term, a term-valued variable, or a unary
    operation.

    Variables are symbols marked with a constraint.  A variable
    "$name" matches any term.  "%name" matches any number.

    There are four unary prefix operators, each of which is a
    pure function from term to term: "=", "+", "*", and ">".
    Their semantics follow.

6.4 Evaluation semantics

    For any term $term, to compute U($term):

        - find the first pattern, in order, that matches $term.

        - substitute its variable matches into its definition.

        - compute the substituted definition.

    Iff this sequence of steps terminates, U($term) "completes."
    Otherwise it "chokes."

    Evaluation is strict: incorrect completion is a bug.  Choking
    is U's only error or exception mechanism.

6.5 Simple operators: equal, increment, evaluate

    =($a $b) is 0 if $a and $b are equal; 1 if they are not.

    +%n is %n plus 1.

    *$a is U($a).

6.6 The follow operator

    >($a $b $c) is always 0.  But it does not always complete.

    We say "$c follows $b in $a" iff, for every $term:

        if *($a 5 $b $term) chokes:
            *($a 5 $c $term) chokes.

        if *($a 5 $b $term) completes:
            either:
                *($a 5 $c $term) completes, and
                *($a 5 $c $term) equals *($a 5 $b $term)
            or:
                *($a 5 $c $term) chokes.

    If $c follows $b in $a, >($a $b $c) is 0.

    If this statement cannot be shown (ie, if there exists any
    $term that falsifies it, generates an infinitely recursive
    series of follow tests, or is inversely self-dependent, ie,
    exhibits Russell's paradox), >($a $b $c) chokes.

7 Implementation issues

    This section is not normative.

7.1 The follow operator

    Of course, no algorithm can completely implement the follow
    operator.  So no program can completely implement U.

    But this does not stop us from stating the correctness of
    a partial implementation - for example, one that assumes
    a hardcoded set of follow cases, and fails when it would
    otherwise have to compute a follow case outside this set.

    U calls this a "trust failure."  One way to standardize trust
    failures would be to standardize a fixed set of follow cases
    as part of the definition of U.  However, this is equivalent
    to standardizing a fixed trusted code base.  The problems
    with this approach are well-known.

    A better design for U implementations is to depend on a
    voluntary, unstandardized failure mechanism.  Because all
    computers have bounded memory, and it is impractical to
    standardize a fixed memory size and allocation strategy,
    every real computing environment has such a mechanism.

    For example, packet loss in an unreliable packet protocol,
    such as UDP, is a voluntary failure mechanism.

    If the packet transfer function of a stateful UDP server is
    defined in terms of U, failure to compute means dropping a
    packet.  If the server has no other I/O, its semantics are
    completely defined by its initial state and packet function.

7.2 Other unstandardized implementation details

    A practical implementation of U will detect and log common
    cases of choking.  It will also need a timeout or some other
    unspecified mechanism to abort undetected infinite loops.

    (Although trust failure, allocation failure or timeout, and
    choke detection all depend on what is presumably a single
    voluntary failure mechanism, they are orthogonal and should
    not be confused.)

    Also, because U is so abstract, differences in implementation
    strategy can result in performance disparities which are
    almost arbitrarily extreme.  The difficulty of standardizing
    performance is well-known.

    No magic bullet can stop these unstandardized issues from
    becoming practical causes of lock-in and incompatibility.
    Systems which depend on U must manage them at every layer.
\end{lstlisting}

\section{The Future of Nock}

SKEW
PLAN

The most likely motivation for a future decrement seems to be the likelihood that some current operation is underspecified, and an implicit choice has been made by every contemporary implementation but should be surfaced.

\section{Conclusion}

Nock 1K should be the final form of the Nock \textsc{isa}, leaving one decrement for Gnon.

\printbibliography
\end{document}